---
title: "error_modeling"
author: "Sai Shreyas Bhavanasi"
date: '2022-04-06'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("visualization_and_analysis_scripts/functions.R")
errs_ = create_error_df()
library(tidyverse)
library(lubridate)
```

Model mean absolute error for a given month for each city on predictors

```{r}
get_rmse <- function(mod, test_data,y){
  
  preds <- predict(mod, newdata = test_data, na.rm = T)
  sqrt(mean((preds - y)^2, na.rm = T))
}
```

```{r}
cities <- read.csv("data/cities.csv")
names(cities) = tolower(names(cities))

# get mainland cities
cities <- cities %>% filter(lon > -150 & lon < -50 & lat > 20 & lat< 50)


cities$city = stringr::str_replace_all(cities$city, "_", " ")

# function returns a dataframe with mean abs error for each city given the input month
get_df <- function(month){
  # get errors for yesterday's high
  errs <- errs_ %>%
    filter(!is.na(error_hi_prev_AM), month(date) == month) %>%
    select(date, error_hi_prev_AM, city)
  unq_cities <- unique(errs$city)
  
  # combine city with error df
  # cities has city predictors
  # errs has city forecases
  new_df <- suppressMessages(dplyr::left_join(errs, cities,  by.x = "city", by.y = "city"))
  
  # want mean absolute error for a given month for each city
  smr_df = suppressMessages(new_df %>% group_by(city, elevation, wind, lat, distance_to_coast,koppen, lon, elevation_change_four,elevation_change_eight, state) %>% 
                              summarize(mean_temp = mean(abs(error_hi_prev_AM))))
  
  #remove na values
  smr_df <- as.data.frame(smr_df)
  smr_df <- smr_df[!is.na(smr_df$elevation_change_four),]
  smr_df <- smr_df[!is.na(smr_df$mean_temp ),]
  
  
  # convert koppen to A,B,C,D,E groups for fewer groups 
  # with more data
  smr_df$kopGrp <- stringr::str_extract(smr_df$koppen, ".")
  
  # remove state, city, and koppen from predictors
  smr_df <- smr_df %>% select_if(!names(.) %in% c("state","city","koppen"))
  
  smr_df
}

smr_df = get_df(2)
# split into test/train
ind_train <- sample(1:nrow(smr_df), .85 * nrow(smr_df))
ind_test <- setdiff(1:nrow(smr_df), ind_train)
dd_train <- smr_df[ind_train,]
dd_test <- smr_df[ind_test,]

```

step AIC
```{r}
mod <- lm(mean_temp ~ . , data = dd_train)
get_rmse(mod,dd_test, dd_test$mean_temp)
summary(mod)
mod_aic <- MASS::stepAIC(mod, direction = "both", trace = 0)
get_rmse(mod_aic,dd_test, dd_test$mean_temp)
summary(mod_aic)
```


Get the error for each month

```{r}


res_df = data.frame()
predictors = list()
for (j in 1:50) {
  ind_train <- sample(1:nrow(smr_df), .85 * nrow(smr_df))
  ind_test <- setdiff(1:nrow(smr_df), ind_train)
  for (i in 1:12) {
    smr_df = get_df(i)
    dd_train <- smr_df[ind_train, ]
    dd_test <- smr_df[ind_test, ]
    mod <- lm(mean_temp ~ . , data = dd_train)
    mod_aic <- MASS::stepAIC(mod, direction = "both", trace = 0)
    res = get_rmse(mod_aic, dd_test, dd_test$mean_temp)
    res_df = rbind(res_df,
                   data.frame(
                     month = i,
                     rmse = res,
                     adj_r2 = summary(mod_aic)$adj.r.squared))
      predictors = append(predictors, list(names(mod_aic$coefficients)))

  }
}

results <- res_df %>% group_by(month) %>% summarize(
  lower_rmse = mean(rmse) - (2*sd(rmse)),
  mean_rmse = mean(rmse),
  upper_rmse = mean(rmse) + (2*sd(rmse)),
  lower_adj_r2 =  mean(adj_r2) - (2*sd(adj_r2)),
  mean_adj_r2 = mean(adj_r2),
  upper_adj_r2 = mean(adj_r2) + (2*sd(adj_r2)))

as.data.frame(results)

mean(results$mean_rmse)
mean(results$mean_adj_r2)

```
```{r}
ggplot()+
  geom_line(aes(y = results$mean_rmse, x = 1:12))+
    geom_line(aes(y = results$upper_rmse, x = 1:12), color = "blue")+
  geom_line(aes(y = results$lower_rmse, x = 1:12), color = "blue")+
  xlim(c(1,12))+
  coord_cartesian(expand = F)+
  xlab("Month")+
  ylab("RMSE")

ggplot()+
  geom_line(aes(y = results$mean_adj_r2, x = 1:12))+
    geom_line(aes(y = results$upper_adj_r2, x = 1:12), color = "blue")+
  geom_line(aes(y = results$lower_adj_r2, x = 1:12), color = "blue")+
  xlim(c(1,12))+
  coord_cartesian(expand = F)+
  xlab("Month")+
  ylab("Adj R2")
```


Frequency of predictors across the months according to stepAIC
                     Var1 Freq
1             (Intercept)   12
9                     lat   11
2       distance_to_coast   10
10                    lon    8
3               elevation    4
6                 kopGrpB    4
7                 kopGrpC    4
8                 kopGrpD    4
11                   wind    4
5   elevation_change_four    3
4  elevation_change_eight    2

Results after running stepaic 50 times.

   month lower_rmse mean_rmse upper_rmse lower_adj_r2
1      1  0.5655403 0.7899490  1.0143578   0.21700663
2      2  0.6330356 0.9970278  1.3610201   0.33582166
3      3  0.3808489 0.5359191  0.6909894   0.33232014
4      4  0.3896248 0.5690769  0.7485290   0.08559992
5      5  0.3772287 0.5322062  0.6871836   0.19552374
6      6  0.3319311 0.4466488  0.5613664   0.07587275
7      7  0.2782489 0.4242117  0.5701745   0.21486350
8      8  0.3392265 0.4917794  0.6443323   0.09976164
9      9  0.2665732 0.4078467  0.5491201   0.19736553
10    10  0.2962184 0.4424104  0.5886025   0.39565044
11    11  0.3694955 0.5575143  0.7455330   0.20216894
12    12  0.3959256 0.6035745  0.8112234   0.17284282
   mean_adj_r2 upper_adj_r2
1    0.2673685    0.3177303
2    0.3839360    0.4320504
3    0.3977495    0.4631788
4    0.1430028    0.2004058
5    0.2557883    0.3160529
6    0.1373210    0.1987692
7    0.2648093    0.3147551
8    0.1604088    0.2210560
9    0.2485308    0.2996961
10   0.4381561    0.4806617
11   0.2650332    0.3278975
12   0.2424824    0.3121220

mean(results$mean_rmse)
[1] 0.5665137
mean(results$mean_adj_r2)
[1] 0.2670489

```{r}
  predictors = c()
  ind_train <- sample(1:nrow(smr_df), .85 * nrow(smr_df))
  ind_test <- setdiff(1:nrow(smr_df), ind_train)
  for (i in 1:12) {
    smr_df = get_df(i)
    dd_train <- smr_df[ind_train, ]
    dd_test <- smr_df[ind_test, ]
    mod <- lm(mean_temp ~ . , data = dd_train)
    mod_aic <- MASS::stepAIC(mod, direction = "both", trace = 0)
    res = get_rmse(mod_aic, dd_test, dd_test$mean_temp)
    res_df = rbind(res_df,
                   data.frame(
                     month = i,
                     rmse = res,
                     adj_r2 = summary(mod_aic)$adj.r.squared))
      predictors = append(predictors, list(names(mod_aic$coefficients)))

  }

preds = c(
"(Intercept)",
"lat"   ,
"distance_to_coast",
"lon",    
"elevation",    
"kopGrpB",    
"kopGrpC",    
"kopGrpD",    
"wind",    
"elevation_change_four",    
"elevation_change_eight")
df = data.frame()
for(i in 1:12){
df = rbind(df, matrix(preds %in% predictors[[i]], ncol = 11 ))
}
names(df) = preds
```


```{r}
cities <- read.csv("data/cities.csv")
names(cities) = tolower(names(cities))

# get mainland cities
cities <- cities %>% filter(lon > -150 & lon < -50 & lat > 20 & lat< 50)


cities$city = stringr::str_replace_all(cities$city, "_", " ")

# function returns a dataframe with mean abs error for each city given the input month

get_df <- function(season){
  # get errors for yesterday's high
  errs <- errs_ %>%
    filter(!is.na(error_hi_prev_AM), month(date) %in% winter ) %>%
    select(date, error_hi_prev_AM, city)
  unq_cities <- unique(errs$city)
  
  # combine city with error df
  # cities has city predictors
  # errs has city forecases
  new_df <- suppressMessages(dplyr::left_join(errs, cities,  by.x = "city", by.y = "city"))
  
  # want mean absolute error for a given month for each city
  smr_df = suppressMessages(new_df %>% group_by(city, elevation, wind, lat, distance_to_coast,koppen, elevation_change_four,elevation_change_eight, state) %>% 
                              summarize(mean_temp = mean(abs(error_hi_prev_AM))))
  
  #remove na values
  smr_df <- as.data.frame(smr_df)
  smr_df <- smr_df[!is.na(smr_df$elevation_change_four),]
  smr_df <- smr_df[!is.na(smr_df$mean_temp ),]
  
  
  # convert koppen to A,B,C,D,E groups for fewer groups 
  # with more data
  smr_df$kopGrp <- stringr::str_extract(smr_df$koppen, ".")
  
  # remove state, city, and koppen from predictors
  smr_df <- smr_df %>% select_if(!names(.) %in% c("state","city","koppen"))
  
  smr_df
}

smr_df = get_df(2)
# split into test/train
ind_train <- sample(1:nrow(smr_df), .85 * nrow(smr_df))
ind_test <- setdiff(1:nrow(smr_df), ind_train)
dd_train <- smr_df[ind_train,]
dd_test <- smr_df[ind_test,]

```


```{r}

res_df = data.frame()
seasons = list( c(3,4,5),c(9,10,11),c(6,7,8),c(1,2))
  for(k in 1:4){
      predictors = c()

    for (j in 1:50) {
  ind_train <- sample(1:nrow(smr_df), .85 * nrow(smr_df))
  ind_test <- setdiff(1:nrow(smr_df), ind_train)
    smr_df = get_df(as.vector(seasons[[k]]))
    dd_train <- smr_df[ind_train, ]
    dd_test <- smr_df[ind_test, ]
    mod <- lm(mean_temp ~ . , data = dd_train)
    mod_aic <- MASS::stepAIC(mod, direction = "both", trace = 0)
    res = get_rmse(mod_aic, dd_test, dd_test$mean_temp)
    # res_df = rbind(res_df,
    #                data.frame(
    #                  rmse = res,
    #                  adj_r2 = summary(mod_aic)$adj.r.squared))
      predictors = append(predictors, list(names(mod_aic$coefficients)))
    }
      temp_df = as.data.frame(table(unlist(predictors))/50)
      temp_df$season = k
    res_df = rbind(res_df, temp_df)
    
}

# mean(res_df[,1]) 
# quantile(res_df[,1], probs = c(.025, .975))
# winter_preds = table(unlist(predictors))/50
```
SPRING
> mean(res_df[,1]) 
[1] 0.3907053
> quantile(res_df[,1], probs = c(.025, .975))
     2.5%     97.5% 
0.2490855 0.5250925 

SUMMER
> mean(res_df[,1]) 
[1] 0.2912825
> quantile(res_df[,1], probs = c(.025, .975))
     2.5%     97.5% 
0.2072269 0.3669929 

FALL
> mean(res_df[,1]) 
[1] 0.3824958
> quantile(res_df[,1], probs = c(.025, .975))
     2.5%     97.5% 
0.2531523 0.5402776 

WINTER
> mean(res_df[,1]) 
[1] 0.5795069
> quantile(res_df[,1], probs = c(.025, .975))
     2.5%     97.5% 
0.4001793 0.7527326 

```{r}
plot.df = data.frame(
  season = c("winter", "spring", "fall", "summer"),
  lower = c(.4,.24,.2,.25),
  upper = c(.75,.52,.36,.54),
  mean = c(.57, .39, .29,.38)
)
plot.df$season = factor(plot.df$season, ordered = T, levels =c( "spring", "summer", "fall","winter") )

ggplot(data = plot.df)+
  geom_point(aes(x = season, y = mean))+
  geom_errorbar(aes(x = season, ymin = lower, ymax = upper))+
  theme_classic()+
  ylab(label= "RMSE")
```

```{r}
ggplot()+
  geom_jitter(data = res_df, aes(x = Var1, y = Freq, color = season), width = 0.1, size = 3)
res_df$season <- as.factor(res_df$season)
levels(res_df$season) <- c("spring","fall","summer", "winter")
```

