---
title: "model_seasons"
author: "Sai Shreyas Bhavanasi"
date: '2022-04-19'
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
```

```{r}
df = read.csv("data/email_data_expanded.csv")
df$city = as.factor(df$city)
df$state = as.factor(df$state)
df$high_or_low = as.factor(df$high_or_low)
df$forecast_outlook = as.factor(df$forecast_outlook)
df$forecast_hours_before = as.factor(df$forecast_hours_before)

df$date = as.Date(df$date)
str(df)
df <- df %>% filter(forecast_hours_before == 24, high_or_low == "high")
df %>% group_by(forecast_outlook) %>% summarise(count = n())
```


```{r}
df <- df %>% filter(city == "ST_LOUIS")
levels(df$forecast_outlook)<- fct_collapse(df$forecast_outlook,
                    SNOW = c("BLGSNO", "BLZZRD","FZDRZL", "FZRAIN","FLRRYS","RNSNOW","SLEET", "SNOSHW"),
                    RAIN = c("CLOUDY", "DRZL","SHWRS","RAIN","TSTRMS" ),
                    CLOUDY = c("MOCLDY", "PTCLDY", "CLOUDY"),
                    NULL = "H")

df <- df %>% mutate(error = observed_temp - forecast_temp)

df %>% group_by(forecast_outlook) %>% summarise(count = n())
```

```{r}
cities <- read.csv("data/cities.csv")
names(cities) = tolower(names(cities))

# get mainland cities
cities <- cities %>% filter(lon > -150 & lon < -50 & lat > 20 & lat< 50)
new_df <- suppressMessages(dplyr::left_join(df, cities,  by.x = "city", by.y = "city"))

new_df = new_df %>% filter((year(date) == 2021 & month(date) %in% 3:12 )|(year(date) == 2022 & month(date) %in% 1:2 ))

```

```{r}
winter = c(12,1,2)
spring = c(3,4,5)
summer = c(6,7,8)
fall = c(9,10,11)
new_df$season <-sapply(month(new_df$date), FUN = function(x)
       if(x %in% spring){"spring"}
       else if(x %in% summer){"summer"}
       else if(x %in% winter){"winter"}
       else if(x %in% fall){"fall"}
  )
new_df$season <- as.factor(new_df$season)
new_df$kopGrp <- stringr::str_extract(new_df$koppen, ".")

```

```{r}
# train = new_df %>% 
#   select(-date, -city, -state, -high_or_low,
#          -forecast_hours_before, -observed_temp, -forecast_temp,
#          -lon, -koppen, -kopGrp)
train = new_df %>% select(season,error,observed_precip, forecast_outlook)
train = train %>% mutate(error = abs(error))
train = na.omit(train)

mod <- lm(error ~ ., data = train)
summary(mod)
names(mod$coefficients)

mod_aic <- MASS::stepAIC(mod, direction = "both", trace = 0)
summary(mod_aic)
```

2021 - march-december [spring,summer,fall,winter]
2022 - jan-feb [winter]
feb has only till feb 1st, need to update data.

outlook - combined all snowy stuff into one.
          rainy stuff into one
          cloudy into one.
          

r2 of 0.05 when considering rain and season
stepAIC removed forecast kept rain and season
only winter was a significant predictor.
precip was also significant.

Error is not dependant on time based on acf plot.
